Constructing Sailing Match Race Schedules: Round Robin Pairing Lists

### 2015-04-07 paper.tex
Spelling

This work describes a round robin competition format that arises from sailing, known as match
racing. Competitors, i.e. {\em skippers}, compete against each other in a series of matches taking
place in rounds, known as {\em flights}.  A match is composed of two skippers, with each skipper in
a boat on his own. Skippers in a match set off together, one on the port side, the other starboard,
and first home wins that match. Criteria published by the Interational Sailing Federation (ISAF)
dictate what makes a legal match race schedule and what should be optimised. This is a rich set of
cirteria, 13 in all, and as far as we know, all published schedules have been produced by hand. This
is a daunting task. A close inspection of these schedules reveals that most are illegal, violating
many of the match race scheduling criteria, many schedules are missing and some that are legal are
far from optimal.

This paper describes the scheduling problem. We present a constraint programming solution and
document how this model was constructed incrementally. We believe this illustrates how adaptable
constraint programming is, i.e. given a fairly robust model this can be incrementally developed to
address all of the required criteria. Our schedules are then constructed in a sequence of stages,
not dissimilar to the hybrid and decomposition based approaches in \cite{lombardi2012}. Some ISAF
published schedules are presented along with some of our own, some of these being new.  These
schedules are esentially blank forms, made available to event organisers, which can be filled in
with names of actual competitors, i.e.\ these schedules are reusable.

@@

Table \ref{10-06} shows the ISAF schedule on the right for 10 skippers and 6 boats and on the right,
our schedule.  The ISAF schedule violates criteria 5 (no more than 2 consequitive port or starboard
assignments) 10 times (twice for skipper 0, 1 and 6 and once for skippers 3, 7, 8 and 9).  Criteria
12 (new boats do not sail first) is violated 7 times (in flight 3 for skippers 6 and 2, flight 5 for
skipper 8, flight 6 for skipper 7, flight 8 for skipper 1, flight 9 for skipper 4 and flight 12 for
skipper 4 again). There are 22 boat changes, with the minimum changes for a skipper being 1 and the
maximum 3. Our schedule, right of Table \ref{10-06} satisfies all criteria, has 12 boat changes with
the minimum changes for a skipper being 0 and the maximum 2, and a maximum imbalance of 1.

>>>>

This work describes a round robin competition format that arises from sailing, known as match
racing. Competitors, i.e. {\em skippers}, compete against each other in a series of matches taking
place in rounds, known as {\em flights}.  A match is composed of two skippers, with each skipper in
a boat on his own. Skippers in a match set off together, one on the port side, the other starboard,
and first home wins that match. Criteria published by the International Sailing Federation (ISAF)
dictate what makes a legal match race schedule and what should be optimised. This is a rich set of
criteria, 13 in all, and as far as we know, all published schedules have been produced by hand. This
is a daunting task. A close inspection of these schedules reveals that most are illegal, violating
many of the match race scheduling criteria, many schedules are missing and some that are legal are
far from optimal.

This paper describes the scheduling problem. We present a constraint programming solution and
document how this model was constructed incrementally. We believe this illustrates how adaptable
constraint programming is, i.e. given a fairly robust model this can be incrementally developed to
address all of the required criteria. Our schedules are then constructed in a sequence of stages,
not dissimilar to the hybrid and decomposition based approaches in \cite{lombardi2012}. Some ISAF
published schedules are presented along with some of our own, some of these being new.  These
schedules are essentially blank forms, made available to event organisers, which can be filled in
with names of actual competitors, i.e.\ these schedules are reusable.

@@

Table \ref{10-06} shows the ISAF schedule on the right for 10 skippers and 6 boats and on the right,
our schedule.  The ISAF schedule violates criteria 5 (no more than 2 consecutive port or starboard
assignments) 10 times (twice for skipper 0, 1 and 6 and once for skippers 3, 7, 8 and 9).  Criteria
12 (new boats do not sail first) is violated 7 times (in flight 3 for skippers 6 and 2, flight 5 for
skipper 8, flight 6 for skipper 7, flight 8 for skipper 1, flight 9 for skipper 4 and flight 12 for
skipper 4 again). There are 22 boat changes, with the minimum changes for a skipper being 1 and the
maximum 3. Our schedule, right of Table \ref{10-06} satisfies all criteria, has 12 boat changes with
the minimum changes for a skipper being 0 and the maximum 2, and a maximum imbalance of 1.
